---
title: "practical machine learning"
author: "fridosj"
date: "15/07/2021"
output: html_document
---
Based on a dataset provide by HAR [http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har) we will try to train a predictive model to predict what exercise was performed using a dataset with 159 features
## Loading libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret); library(rattle); library(randomForest); library(corrplot); library(rpart)
set.seed(12345)
```
## Data Loading and creating train-test datasets
The next step is loading the dataset from the URL provided above. The training dataset is then partinioned in 2 to create a Training set (70% of the data) for the modeling process and a Test set (with the remaining 30%) for the validations. The testing dataset is not changed and will only be used for the quiz results generation.
UrlTrain <- [ttp://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](train data)
UrlTest  <- [http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](test data)


```{r}
training <- read.csv(file = "C:/Users/HP/Downloads/pml-training.csv")
testing  <- read.csv(file="C:/Users/HP/Downloads/pml-testing.csv")
inTrain  <- createDataPartition(training$classe, p=0.7, list=FALSE)
TrainSet <- training[inTrain, ]
TestSet  <- training[-inTrain, ]
dim(TrainSet)
```
Both created datasets have 160 variables. Those variables have plenty of NA, that can be removed with the cleaning procedures below. The Near Zero variance (NZV) variables are also removed and the ID variables as well.

```{r}
NZV <- nearZeroVar(TrainSet)
TrainSet <- TrainSet[, -NZV]
TestSet  <- TestSet[, -NZV]
AllNA    <- sapply(TrainSet, function(x) mean(is.na(x))) > 0.95
TrainSet <- TrainSet[, AllNA==FALSE]
TestSet  <- TestSet[, AllNA==FALSE]
TrainSet <- TrainSet[, -(1:5)]
TestSet  <- TestSet[, -(1:5)]
dim(TrainSet)
```

```{r}
dim(TestSet)
```
## Correlation anlaysis
A correlation among variables is analysed before proceeding to the modeling procedures.
```{r}
corMatrix <- cor(TrainSet[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower",tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```
The highly correlated variables are shown in dark colors in the graph above. To make an evem more compact analysis, a PCA (Principal Components Analysis) could be performed as pre-processing step to the datasets. Nevertheless, as the correlations are quite few, this step will not be applied for this assignment.
## Prediction Model Building
Three methods will be applied to model the regressions (in the Train dataset) and the best one (with higher accuracy when applied to the Test dataset) will be used for the quiz predictions. The methods are: Random Forests, Decision Tree and Generalized Boosted Model, as described below. A Confusion Matrix is plotted at the end of each analysis to better visualize the accuracy of the models.

### a) Random Forest
```{r}
set.seed(12345)
myModelFilename <- "rf.RData"
if (!file.exists(myModelFilename)) {
  
  # Parallel cores  
  library(doParallel)
  ncores <- makeCluster(detectCores() - 1)
  registerDoParallel(cores=ncores)
  getDoParWorkers() # 3    
  
  # use Random Forest method with Cross Validation, 4 folds
  myModel <- train(classe ~ .
                   , data = TrainSet
                   , method = "rf"
                   , metric = "Accuracy"  # categorical outcome variable so choose accuracy
                   , preProcess=c("center", "scale") # attempt to improve accuracy by normalising
                   , trControl=trainControl(method = "cv"
                                            , number = 4 # folds of the training data
                                            , p= 0.60
                                            , allowParallel = TRUE 
                                            #                                       , seeds=NA # don't let workers set seed 
                   )
  )
  
  save(myModel, file = "myModel.RData")
  # 3:42 .. 3:49 without preProcess
  # 3:51 .. 3:58 with preProcess
  stopCluster(ncores)
} else {
  # Use cached model  
  load(file = myModelFilename, verbose = TRUE)
}
print(myModel, digits=4)
```



### b) Decision trees
```{r}
set.seed(12345)
modfittree <- rpart(classe ~ ., data=TrainSet, method="class")
fancyRpartPlot(modfittree)
```

```{r}
predtree <- predict(modfittree, newdata=TestSet)
print(predtree,digits=4)
```

### c) Generalized Boosted Model
```{r}
set.seed(12345)
controlGBM <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
modFitGBM  <- train(classe ~ ., data=TrainSet, method = "gbm",
                    trControl = controlGBM, verbose = FALSE)
modFitGBM$finalModel

```

```{r}
predictGBM <- predict(modFitGBM, newdata=TestSet)
print(modFitGBM,digits=4)
```


## Applying the Selected Model to the Test Data
The accuracy of the 3 regression modeling methods above are:

Random Forest : 0.9972 Decision Tree : 0.7368 GBM : 0.9839 In that case, the Random Forest model will be applied to predict the 20 quiz results (testing dataset) as shown below.


```{r}
predictTEST <- predict(myModel, newdata=testing)
predictTEST
```
